---
title: "Data Science Avancées"
# output: pdf_document
date: "2023-03-28"
output:
  pdf_document: 
    fig_height: 6
    keep_tex: yes
    highlight: kate
    latex_engine: xelatex
    number_sections: yes
  html_document:
    highlight: kate
  word_document:
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Importation du jeu de données
```{r}
bbc <- read.table(
  file= "~/Documents/Uparis/M1MLSD2223/ue_data2/text-clustering/core/datasets/data/bbc.csv",
  header=TRUE,
  sep=",",
  fileEncoding= "latin1")
```

```{r}
# on considère indexation de base 
df.bbc = dplyr::tibble(bbc[-1])
head(df.bbc)
```

```{r}
# convert as factor
df.bbc = df.bbc |>
  dplyr::mutate(label = factor(label))
```

```{r}
df.bbc.docs = dplyr::tibble(df.bbc["text"])
df.bbc.labels = dplyr::tibble(df.bbc["label"])
```

```{r}
str(df.bbc.labels)
```



```{r}
# get docs frame as vector of docs
df.bbc.docs.as.vec = dplyr::pull(df.bbc.docs, text)
df.bbc.labels.as.vec = dplyr::pull(df.bbc.labels, label)
```

# Tokenizer
```{r}
clust_vec = textTinyR::tokenize_transform_vec_docs(
  object = df.bbc.docs.as.vec, 
  as_token = T,
  to_lower = T, 
  remove_punctuation_vector = F,
  remove_numbers = F, 
  trim_token = T,
  split_string = T,
  split_separator = " \r\n\t.,;:()?!//", 
  remove_stopwords = T,
  language = "english", 
  min_num_char = 3, 
  max_num_char = 100,
  stemmer = "porter2_stemmer", 
  threads = 4,
  verbose = T)
```

I'll build also the term matrix as I'll need the global-term-weights

```{r}
utl = textTinyR::sparse_term_matrix$new(
  vector_data = df.bbc.docs.as.vec, 
  file_data = NULL,
  document_term_matrix = TRUE)

# term-matrix
tm = utl$Term_Matrix(
  sort_terms = FALSE, 
  to_lower = T, 
  remove_punctuation_vector = F,
  remove_numbers = F, trim_token = T, 
  split_string = T, stemmer = "porter2_stemmer",
  split_separator = " \r\n\t.,;:()?!//", 
  remove_stopwords = T, language = "english", 
  min_num_char = 3, max_num_char = 100,
  print_every_rows = 100000, normalize = NULL, 
  tf_idf = F, threads = 6, verbose = T)

gl_term_w = utl$global_term_weights()
str(gl_term_w)
```


# build Doc2vec matrix using token

Estimate word vectors dimension

```{r}
textTinyR::dims_of_word_vecs(
  input_file = "/home/godwin/Documents/Uparis/M1MLSD2223/ue_data2/tmp/glove.840B.300d.txt", 
  read_delimiter = "\n")
```

Let's build doc2vec matrix from word embbeded

```{r}
# class init
init = textTinyR::Doc2Vec$new(
  token_list = clust_vec$token, 
  word_vector_FILE = "/home/godwin/Documents/Uparis/M1MLSD2223/ue_data2/tmp/glove.840B.300d.txt",
  print_every_rows = 5000, 
  verbose = TRUE, 
  copy_data = T)  # if true we can inspect 
```

```{r}
#res_wv = init$pre_processed_wv()                           
# 
#str(res_wv)
```



```{r}
doc2_idf = init$doc2vec_methods(
  method = "idf", 
  global_term_weights = gl_term_w, 
  threads = 6)

doc2_sum = init$doc2vec_methods(method = "sum_sqrt", threads = 6)
doc2_norm = init$doc2vec_methods(method = "min_max_norm", threads = 6)
```

```{r}
rows_cols = 1:5
doc2_idf[rows_cols, rows_cols]
```



```{r}
dim(doc2_idf)
```


# Reduced k-means

```{r}
# apply Reduced K-means to iris (class variable is excluded)
out.cluspca = clustrd::cluspca(
  doc2_idf,  5, 2, seed = 1234)
summary(out.cluspca)
```

An important feature of Reduced K-means is that we can visualize observations, cluster centers and variables on a low-dimensional space, similar to PCA.

```{r}
# plotting the RKM solution
plot(out.cluspca, cludesc = TRUE)
```

It is helpful to compute the confusion matrix between the true cluster partition (df_bbc[, 5]) and the one obtained by Reduced K-means
(out.cluspca$cluster). The agreement between the obtained partition and the true partition is moderate, according to an adjusted Rand index of .62.

```{r}
# use labels as vector and compare it to vectors of predicted labels (cluster)
table(df.bbc.labels.as.vec, out.cluspca$cluster)
```

# Choosing the number of clusters and dimensions

```{r}

scal_dat = ClusterR::center_scale(doc2_sum)     # center and scale the data

# Cluster quality assessment based on the average silhouette width 
# in the low dimensional space
bestRKM = clustrd:: tuneclus(scal_dat, 3:8, 2:4, method = "RKM", criterion = "asw",
          dst = "low")
bestRKM
```




